Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-02-16 csf-2017/paper.tex
pass over example section

From another point of view, the Fundamental Theorem also serves as a sanity
check of the logical relation as it guarantees that a large class of
capabilities actually inhabits it.

@@

\subsection{Encapsulation of local state}
\texttt{\footnotesize{f1}} and \texttt{\footnotesize{f2}} in
\figurename~\ref{fig:prog-f1-and-f2} demonstrate the capability machine's
encapsulation of local state. They are very similar: both store some local
state, call an untrusted piece of code ($\adv$), and then test whether the local
state is unchanged. They differ in the way they do this. Program
\texttt{\footnotesize{f1}} assumes access to a stack pointer and uses the stack
to store its local state and the stack-based calling convention
\texttt{\footnotesize{scall}} to call the adversary. On the other hand,
\texttt{\footnotesize{f2}} uses malloc to allocate memory for its local state
and uses an activation-record based calling convention (described in the
technical appendix) to run the adversarial code.

@@

In order to formulate results about \texttt{\footnotesize{f1}} and
\texttt{\footnotesize{f2}}, we need a way to observe whether the assertion
fails. To this end, we assume they have access to a flag (an address in memory)
that they set to $1$ if the assertion fails and then halt. The correctness lemma
for \texttt{\footnotesize{f1}} then states:
\begin{lemma}
  \label{lem:correctness-f1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_\adv & \;\defeq ((\entry,\glob),\dots) & c_{\var{stk}} & \;\defeq ((\rwlx,\local),\dots)\\
    c_{f1} & \;\defeq ((\rwx,\glob),\dots) & c_\link &\; \defeq ((\readonly,\glob),\dots)\\
    c_\malloc &\; \defeq ((\entry,\glob),\dots) & \reg& \; \in \Regs \\
    m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus \ms_{\malloc} \uplus} \\
      & \multicolumn{3}{@{}l}{\phantom{\;\defeq \;}  \ms_{\var{stk}} \uplus \ms_{\var{frame}}} \\
    \end{array}
\]
where each of the capabilities have an appropriate range of authority and
pointer\footnote{These assumptions are kept intentionally vague for brevity.
  Full statements are in the technical appendix~\citep{technical_appendix}.}.
Furthermore
  \begin{itemize}
  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{\footnotesize{f1}}
  \item $\ms_\flag(\flag) = 0$
  \item $\ms_{\var{link}}$ contains $c_\adv$ and $c_\malloc$
  \item $\hs_\adv$ contains $c_\link$ and otherwise only instructions.
  \end{itemize}
  If $(\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step^* (\halted,m')$,
  then $m'(\flag) = 0$
\end{lemma}

@@

The final example, \texttt{\footnotesize{g1}} with \texttt{\footnotesize{f4}},
is a faithful translation of a tricky example known from the
literature~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists of two
parts, \texttt{\footnotesize{g1}} and \texttt{\footnotesize{f4}}.
\texttt{\footnotesize{g1}} is a closure generator that generates closures with
one variable $x$ set to $0$ in its environment and \texttt{\footnotesize{f4}} as
the program (note that we can omit some calling convention security measures
because the stack is not used or passed to the adversary in the closure
generator). \texttt{\footnotesize{f4}} expects one argument, a callback. It sets
$x$ to $0$ and calls the callback. When it returns, it sets $x$ to $1$ and calls
the callback a second time. When it returns again, it asserts that $x$ is $1$
and returns. This example is more complicated than the previous ones because it
involves a closure invoked by the adversary and an adversary callback invoked by
us. As explained in Section~\ref{sec:stack-and-return-pointer}, this means we
need to check (1) that the stack pointer that the closure receives from the
adversary (it must have the correct permission and fit with our calling
convention) and (2) that the adversary callback is global.

@@

In Lemma~\ref{lem:correctness-g1} the order of control has been inverted
compared to the previous lemmas. In this lemma, the adversary assumes control
first with a capability for the closure creator \texttt{\footnotesize{g1}}. This means we need
to check that all arguments are safe to use and that we clean up before
returning in the end. The inversion of control poses an interesting challenge
when it comes to reasoning about the adversary's local state during the
execution of \texttt{\footnotesize{f4}}. During the execution of \texttt{\footnotesize{f4}} and the callbacks,
the adversary should not be able to rely on its local state from before the call
of \texttt{\footnotesize{f4}}. This is easily done by revoking all the temporary regions of the
world given at the start of \texttt{\footnotesize{f4}}. However, when \texttt{\footnotesize{f4}} returns, the
adversary is again allowed to rely on its old local state which means that we
need to guarantee that the local state is unchanged. This is important because
the return pointer that \texttt{\footnotesize{f4}} receives from the adversary may be local,
and the adversary is allowed to allocate the activation record on the stack
(just like we do) so that they can store and recover his old stack pointer after
\texttt{\footnotesize{f4}} returns.

>>>>

From another point of view, the Fundamental Theorem also serves as a sanity
check of the logical relation as it guarantees a large class of
capabilities actually inhabits it.

@@

\subsection{Encapsulation of local state}
\texttt{\footnotesize{f1}} and \texttt{\footnotesize{f2}} in
\figurename~\ref{fig:prog-f1-and-f2} demonstrate the capability
machine's encapsulation of local state. They are very similar: both
store some local state, call an untrusted piece of code ($\adv$), and
then test whether the local state is unchanged. They differ in the way
they do this. Program \texttt{\footnotesize{f1}} uses our stack-based
calling convention (captured by \texttt{\footnotesize{scall}}) to call
the adversary, so it can use the available stack to store its local
state.  On the other hand, \texttt{\footnotesize{f2}} uses malloc to
allocate memory for its local state and uses an activation-record
based calling convention (described in the technical appendix) to run
the adversarial code.

@@

In order to formulate results about \texttt{\footnotesize{f1}} and
\texttt{\footnotesize{f2}}, we need a way to observe whether the assertion
fails. To this end, we assume they have access to a flag (an address in memory).
If the assertion fails, then the flag is set to $1$ and execution halts. The correctness lemma
for \texttt{\footnotesize{f1}} then states:
\begin{lemma}
  \label{lem:correctness-f1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_\adv & \;\defeq ((\entry,\glob),\dots) & c_{\var{stk}} & \;\defeq ((\rwlx,\local),\dots)\\
    c_{f1} & \;\defeq ((\rwx,\glob),\dots) & c_\link &\; \defeq ((\readonly,\glob),\dots)\\
    c_\malloc &\; \defeq ((\entry,\glob),\dots) & \reg& \; \in \Regs \\
    m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus \ms_{\malloc} \uplus} \\
      & \multicolumn{3}{@{}l}{\phantom{\;\defeq \;}  \ms_{\var{stk}} \uplus \ms_{\var{frame}}} \\
    \end{array}
\]
where each of the capabilities have an appropriate range of authority and
pointer\footnote{These assumptions are kept intentionally vague for brevity.
  Full statements are in the technical appendix~\citep{technical_appendix}.}.
Furthermore
  \begin{itemize}
  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{\footnotesize{f1}}
  \item $\ms_\flag(\flag) = 0$
  \item $\ms_{\var{link}}$ contains $c_\adv$ and $c_\malloc$
  \item $\hs_\adv$ contains $c_\link$ and otherwise only instructions.
  \end{itemize}
  If $(\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step^* (\halted,m')$,
  then $m'(\flag) = 0$
\end{lemma}

@@

The final example, \texttt{\footnotesize{g1}} with \texttt{\footnotesize{f4}},
is a faithful translation of a tricky example known from the
literature~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists of two
parts, \texttt{\footnotesize{g1}} and \texttt{\footnotesize{f4}}.
\texttt{\footnotesize{g1}} is a closure generator that generates closures with
one variable $x$ set to $0$ in its environment and \texttt{\footnotesize{f4}} as
the program (note we can omit some calling convention security measures
because the stack is not used or passed to the adversary in the closure
generator). \texttt{\footnotesize{f4}} expects one argument, a callback. It sets
$x$ to $0$ and calls the callback. When it returns, it sets $x$ to $1$ and calls
the callback a second time. When it returns again, it asserts $x$ is $1$
and returns. This example is more complicated than the previous ones because it
involves a closure invoked by the adversary and an adversary callback invoked by
us. As explained in Section~\ref{sec:stack-and-return-pointer}, this means we
need to check (1) the stack pointer that the closure receives from the
adversary (it must have the correct permission and fit with our calling
convention) and (2) the adversary callback is global.

@@

In Lemma~\ref{lem:correctness-g1} the order of control has been inverted
compared to the previous lemmas. In this lemma, the adversary assumes control
first with a capability for the closure creator \texttt{\footnotesize{g1}}. This means we need
to check that all arguments are safe to use and that we clean up before
returning in the end. The inversion of control poses an interesting challenge
when it comes to reasoning about the adversary's local state during the
execution of \texttt{\footnotesize{f4}}. During the execution of \texttt{\footnotesize{f4}} and the callbacks,
the adversary should not be able to rely on its local state from before the call
of \texttt{\footnotesize{f4}}. This is easily done by revoking all the temporary regions of the
world given at the start of \texttt{\footnotesize{f4}}. However, when \texttt{\footnotesize{f4}} returns, the
adversary is again allowed to rely on its old local state which means we
need to guarantee that the local state is unchanged. This is important because
the return pointer that \texttt{\footnotesize{f4}} receives from the adversary may be local,
and the adversary is allowed to allocate the activation record on the stack
(just like we do) so they can store and recover their old stack pointer after
\texttt{\footnotesize{f4}} returns.
